A generalized model for the charge transport properties of dressed quantum Hall systems

### 2021-08-22 paper/Submission/dressed_quantum_hall/dressed_quantum_hall.tex
Re-structure

%===============================================================================
% Authors of the manuscript
%===============================================================================
\author{Kosala Herath,}
\email[]{kosala.herath@monash.edu}
\author{Malin Premaratne}
\email[]{malin.premaratne@monash.edu}
\affiliation{Advanced Computing and Simulation Laboratory (A$\chi$L), Department of Electrical and Computer Systems Engineering, Monash University, Clayton, Victoria 3800, Australia.}
%===============================================================================

@@

%===============================================================================
% Section 01 : Introduction
%===============================================================================
\section{\label{sec_introduction} Introduction}

@@

The paper is organized as follows. Sec.  \ref{sec_schrodinger_problem} introduces our dressed quantum Hall system and the exact wave function solutions for the given configuration. Sec. \ref{sec_floquet_theory} provides the Floquet theory interpretation of these wave functions.
We introduce Floquet-Fermi golden rule for a quantum Hall system in Sec. \ref{sec_inverse_scattering_time}, and use it in Sec. \ref{sec_floquet_drude_conductivity} to derive analytical expressions for longitudinal components of conductivity.
The derived theoretical model is further analyzed numerically using empirical system parameters and compared with previous studies in Sec. \ref{sec_manipulate_conductivity}.
In Sec. \ref{sec_conclusions}, we summarize our analysis and present our conclusions.
%===============================================================================

%===============================================================================
% Section 02 : Schrodinger problem for a dressed quantum Hall system
%===============================================================================
\section{\label{sec_schrodinger_problem} Schrodinger Problem for a Dressed Quantum Hall System}

Our system consists of a 2DEG placed on the $xy$-plane of the three-dimensional coordinate space. In our analysis, the 2DEG is subjected to a nonoscillating magnetic field $\vb{B} = (0,0,B)^{T}$, which is pointed towards the $z$-axis. In addition, a linearly polarized strong light is applied perpendicular to the 2DEG surface. We specially select the frequency of the dressing field $\omega$ to be in the off-resonant regime such that the field behaves as a purely dressing field. Furthermore, without limiting the generality, we choose a $y$-polarized electric field $\vb{E} = (0,E\sin(\omega t),0)^{T}$ for the linearly polarized dressing field as given in Fig.~\ref{fig:1}.
\begin{figure}[b]
\includegraphics[scale=0.9]{fig_1}
\caption{\label{fig:1} Our 2DEG system only confined in the $xy$-plane while both of the stationary magnetic field $\vb{B}$ and the dressing field are applied perpendicular to the plane of 2DEG. The dressing field is linearly polarized with a $y$-polarized electric field $\vb{E}$.}
\end{figure}
Here $B$ and $E$ represent the amplitudes of the stationary magnetic field and oscillating electric field, respectively.

@@

%===============================================================================
% Section 03 : Floquet theory perspective
%===============================================================================
\section{\label{sec_floquet_theory} Floquet Theory Perspective}

@@

%===============================================================================
% Section 04 : Inverse scattering time analysis
%===============================================================================
\section{\label{sec_inverse_scattering_time}  Inverse Scattering Time Analysis}

@@

\begin{figure}[t]
\includegraphics[scale=0.55]{fig_2}
\caption{\label{fig_3} The dependence of normalized scattering-induced broadening $\Lambda_N$ for each Landau level ($N =0,1,2,3,4$) against $x$-directional momentum value $k_x$ in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a  frequency of $\omega =\SI{2e12}{\radian\per\second}$ and intensity $I =\SI{100}{\watt\per\square\centi\metre}$.
In this calculation, we have assumed that the natural  broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\end{figure}

\begin{figure}[t]
\includegraphics[scale=0.54]{fig_3}
\caption{\label{fig_4} The dependence of normalized scattering-induced broadening $\Lambda_N$ for each Landau level ($N =0,1,2,3,4$) against dressing field intensity $I$, in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a frequency of $\omega =\SI{2e12}{\radian\per\second}$. In this calculation, we have assumed that the natural broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\end{figure}

@@

%===============================================================================
% Section 05 : Floquet-Drude Conductivity in a Dressed Quantum Hall System
%===============================================================================
\section{\label{sec_floquet_drude_conductivity} Floquet-Drude Conductivity in a Dressed Quantum Hall System}

@@

%===============================================================================
% Section 06 : Manipulate Conductivity in Quantum Hall Systems
%===============================================================================
\section{\label{sec_manipulate_conductivity} Manipulate Conductivity in Quantum Hall Systems}

@@

\begin{figure}[t]
\includegraphics[scale=0.54]{fig_4}
\caption{ Normalized longitudinal conductivity $\widetilde{\sigma}^{xx}$ against Fermi level $X_F$ with different intensities $I$ of the external dressing field in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a  frequency of $\omega =\SI{2e12}{\radian\per\second}$ and $I_0 =\SI{100}{\watt\per\square\centi\metre}$. In this calculation, we have assumed that the natural  broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\label{fig:5}
\end{figure}

\begin{figure}[t]
\includegraphics[scale=0.54]{fig_5}
\caption{$3$rd Landau level’s normalized longitudinal conductivity $\widetilde{\sigma}^{xx}$ against Fermi level $X_F$ with different intensities $I$ of the external dressing field in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a frequency of $\omega =\SI{2e12}{\radian\per\second}$ and $I_0 =\SI{100}{\watt\per\square\centi\metre}$.
In this calculation, we have assumed that the natural  broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\label{fig:6}
\end{figure}
%===============================================================================

%===============================================================================
% Section 07 : Conclusions
%===============================================================================
\section{\label{sec_conclusions} Conclusions}

@@

%===============================================================================
% Add appendixes
%===============================================================================
\begin{appendix}

@@

Next, we consider a scattering phenomenon when a electron scatters from a known $t$-$t'$-Floquet state $\ket{\psi_{\beta}(t,t')} $ into a distinct $t$-$t'$-Floquet state $\ket{\Psi_{\alpha}(t,t')}$ with a constant quansienergy $\varepsilon$ as follows
\begin{equation} \label{eq:c15}
  \begin{aligned}
  \ket{\psi_{\beta}(t,t')} &= \exp(-\frac{i}{\hbar}\varepsilon_{\beta} t)
  \ket{\phi_{\beta}(t')} \\
  &
  \xRightarrow{\text{scattering}}
  \ket{\Psi_{\alpha}(t,t')} = \exp(-\frac{i}{\hbar}\varepsilon t)
  \ket{\Phi_{\alpha}(t')}.
  \end{aligned}
\end{equation}
\begin{figure}[b]
  \includegraphics[scale=1.0]{fig_6.pdf}
  \caption{Scattering from known $\ket{\psi_{\beta}(t,t')}$ state to a constant energy state $\ket{\Psi_{\alpha}(t,t')}$ due to the scattering potential created by impurities in the material.}
  \label{fig:2}
\end{figure}
This phenomenon is illustrated in Fig.~\ref{fig:2}.
We can calculate the scattering amplitude for this scattering scenario using the expression derived in Eq.~(\ref{eq:c14}) as follows
\begin{equation} \label{eq:c16}
  \begin{aligned}
    a_{\alpha\beta}(t,t') =
    -\frac{i}{\hbar}
    \int_{0}^t
    e^{\frac{i}{\hbar}\left(\varepsilon_{\beta} - \varepsilon \right)t_1}
    \bra{\phi_{\beta}(t')}
    V(\vb{r}) \ket{\phi_{\alpha}(t')}  dt_1.
  \end{aligned}
\end{equation}
By assuming for a long duration ($t \rightarrow \infty$), we can turn this integral into a delta distribution
\begin{equation} \label{eq:c17}
  \begin{aligned}
    a_{\alpha\beta}(t') =
    -2\pi i \delta(\varepsilon_{\beta} - \varepsilon)Q,
  \end{aligned}
\end{equation}
where $Q = \bra{\phi_{\beta}(t')} V(\vb{r}) \ket{\phi_{\alpha}(t')}$ and using the completeness properties we can re-write this as
\begin{equation} \label{eq:c18}
    Q =
    \sum_{\vb{k}}\sum_{\vb{k'}}
    \braket{\phi_{\beta}(t')}{\vb{k'}}
    \mel{\vb{k'}}{V(\vb{r})}{\vb{k}}
    \braket{\vb{k}}{\phi_{\alpha}(t')}.
\end{equation}
Moreover, separating $x$-directional and $y$-directional momentum components, we can simply this to obtain
\begin{equation} \label{eq:c19}
    Q =
    \sum_{k_x}\sum_{{k'}_x}
    \int_{-\infty}^{\infty} \int_{-\infty}^{\infty}
    V_{\vb{k'},\vb{k}}
    \phi_{\beta}^{\dagger}(\vb{k'},t')
    \phi_{\alpha}(\vb{k},t')  dk_y d{k'}_y,
\end{equation}
with $V_{\vb{k'},\vb{k}} = \mel{\vb{k'}}{V(\vb{r})}{\vb{k}}$.

@@

Moreover, we can identify the $y$-directional current operator component as
\begin{equation} \label{eq:d12}
  \hat{j}_y = \frac{1}{\widetilde{m}} \left(-i\hbar\pdv{y} - \frac{eE}{\omega}\cos(\omega t) \right).
\end{equation}
Using this operator, we can represent the matrix elements of $y$-directional current operator in Floquet mode basis as
\begin{equation} \label{eq:d13}
  \left({j}_y \right)_{nm,n'm'} =
  \mel{n,k_x}{\;\frac{-1}{\widetilde{m}} \left(i\hbar\pdv{y} + \frac{eE}{\omega}\cos(\omega t) \right)}{n',k'_x}.
\end{equation}
After following the same steps done for the $x$-directional current operator, we can identify the $0$-th component of matrix elements for $y$-directional electric current operator as
\begin{equation} \label{eq:d14}
  \begin{aligned}
    \Big({j}^y_{s=0}\Big)_{nm,n'm'}^{electron} = &
    \frac{ie\hbar}{{m_e}}
    \delta_{k_x,k'_x} \\
    & \times
    \left(
    \sqrt{\frac{n}{2}} \delta_{n',n-1}
    - \sqrt{\frac{n+1}{2}} \delta_{n',n+1}
    \right).
  \end{aligned}
\end{equation}
%===============================================================================
\end{appendix}

>>>>

%===============================================================================
% Authors of the manuscript
%===============================================================================
\author{Kosala Herath,}
\email{kosala.herath@monash.edu}
\author{Malin Premaratne}
\email{malin.premaratne@monash.edu}
\affiliation{Advanced Computing and Simulation Laboratory (A$\chi$L), Department of Electrical and Computer Systems Engineering, Monash University, Clayton, Victoria 3800, Australia.}
%===============================================================================

@@

%===============================================================================
% Section 01 : Introduction
%===============================================================================
\section{\label{sec:introduction} Introduction}

@@

The paper is organized as follows. Sec.  \ref{sec:schrodinger_problem} introduces our dressed quantum Hall system and the exact wave function solutions for the given configuration. Sec. \ref{sec:floquet_theory} provides the Floquet theory interpretation of these wave functions.
We introduce Floquet-Fermi golden rule for a quantum Hall system in Sec. \ref{sec:inverse_scattering_time}, and use it in Sec. \ref{sec:floquet_drude_conductivity} to derive analytical expressions for longitudinal components of conductivity.
The derived theoretical model is further analyzed numerically using empirical system parameters and compared with previous studies in Sec. \ref{sec:manipulate_conductivity}.
In Sec. \ref{sec:conclusions}, we summarize our analysis and present our conclusions.
%===============================================================================

%===============================================================================
% Section 02 : Schrodinger problem for a dressed quantum Hall system
%===============================================================================
\section{\label{sec:schrodinger_problem} Schrodinger Problem for a Dressed Quantum Hall System}

Our system consists of a 2DEG placed on the $xy$-plane of the three-dimensional coordinate space. In our analysis, the 2DEG is subjected to a nonoscillating magnetic field $\vb{B} = (0,0,B)^{T}$, which is pointed towards the $z$-axis. In addition, a linearly polarized strong light is applied perpendicular to the 2DEG surface. We specially select the frequency of the dressing field $\omega$ to be in the off-resonant regime such that the field behaves as a purely dressing field. Furthermore, without limiting the generality, we choose a $y$-polarized electric field $\vb{E} = (0,E\sin(\omega t),0)^{T}$ for the linearly polarized dressing field as given in Fig.~\ref{fig:1}.
\begin{figure}[b]
\includegraphics[scale=0.9]{fig_1_system_diagram}
\caption{\label{fig:1} Our 2DEG system only confined in the $xy$-plane while both of the stationary magnetic field $\vb{B}$ and the dressing field are applied perpendicular to the plane of 2DEG. The dressing field is linearly polarized with a $y$-polarized electric field $\vb{E}$.}
\end{figure}
Here $B$ and $E$ represent the amplitudes of the stationary magnetic field and oscillating electric field, respectively.

@@

%===============================================================================
% Section 03 : Floquet theory perspective
%===============================================================================
\section{\label{sec:floquet_theory} Floquet Theory Perspective}

@@

%===============================================================================
% Section 04 : Inverse scattering time analysis
%===============================================================================
\section{\label{sec:inverse_scattering_time}  Inverse Scattering Time Analysis}

@@

\begin{figure}[t]
\includegraphics[scale=0.55]{fig_2_normalized_broadening_with_momentum}
\caption{\label{fig_3} The dependence of normalized scattering-induced broadening $\Lambda_N$ for each Landau level ($N =0,1,2,3,4$) against $x$-directional momentum value $k_x$ in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a  frequency of $\omega =\SI{2e12}{\radian\per\second}$ and intensity $I =\SI{100}{\watt\per\square\centi\metre}$.
In this calculation, we have assumed that the natural  broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\end{figure}

\begin{figure}[t]
\includegraphics[scale=0.54]{fig_3_normalized_broadening_with_intensity}
\caption{\label{fig_4} The dependence of normalized scattering-induced broadening $\Lambda_N$ for each Landau level ($N =0,1,2,3,4$) against dressing field intensity $I$, in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a frequency of $\omega =\SI{2e12}{\radian\per\second}$. In this calculation, we have assumed that the natural broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\end{figure}

@@

%===============================================================================
% Section 05 : Floquet-Drude Conductivity in a Dressed Quantum Hall System
%===============================================================================
\section{\label{sec:floquet_drude_conductivity} Floquet-Drude Conductivity in a Dressed Quantum Hall System}

@@

%===============================================================================
% Section 06 : Manipulate Conductivity in Quantum Hall Systems
%===============================================================================
\section{\label{sec:manipulate_conductivity} Manipulate Conductivity in Quantum Hall Systems}

@@

\begin{figure}[t]
\includegraphics[scale=0.54]{fig_4_conductivity_with_fermi_level}
\caption{ Normalized longitudinal conductivity $\widetilde{\sigma}^{xx}$ against Fermi level $X_F$ with different intensities $I$ of the external dressing field in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a  frequency of $\omega =\SI{2e12}{\radian\per\second}$ and $I_0 =\SI{100}{\watt\per\square\centi\metre}$. In this calculation, we have assumed that the natural  broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\label{fig:5}
\end{figure}

\begin{figure}[t]
\includegraphics[scale=0.54]{fig_5_conductivity_with_fermi_level_zoomed}
\caption{$3$rd Landau level’s normalized longitudinal conductivity $\widetilde{\sigma}^{xx}$ against Fermi level $X_F$ with different intensities $I$ of the external dressing field in a GaAs-based quantum well under a nonoscillating magnetic field with $B = \SI{1.2}{\tesla}$, dressing field with a frequency of $\omega =\SI{2e12}{\radian\per\second}$ and $I_0 =\SI{100}{\watt\per\square\centi\metre}$.
In this calculation, we have assumed that the natural  broadening of $0$-th Landau level $\Gamma_0$ is $\SI{0.24}{\milli\eV}$.}
\label{fig:6}
\end{figure}
%===============================================================================

%===============================================================================
% Section 07 : Conclusions
%===============================================================================
\section{\label{sec:conclusions} Conclusions}

@@

%===============================================================================
% Add appendixes
%===============================================================================
\appendix

@@

Next, we consider a scattering phenomenon when a electron scatters from a known $t$-$t'$-Floquet state $\ket{\psi_{\beta}(t,t')} $ into a distinct $t$-$t'$-Floquet state $\ket{\Psi_{\alpha}(t,t')}$ with a constant quansienergy $\varepsilon$ as follows
\begin{equation} \label{eq:c15}
  \begin{aligned}
  \ket{\psi_{\beta}(t,t')} &= \exp(-\frac{i}{\hbar}\varepsilon_{\beta} t)
  \ket{\phi_{\beta}(t')} \\
  &
  \xRightarrow{\text{scattering}}
  \ket{\Psi_{\alpha}(t,t')} = \exp(-\frac{i}{\hbar}\varepsilon t)
  \ket{\Phi_{\alpha}(t')}.
  \end{aligned}
\end{equation}
\begin{figure}[b]
  \includegraphics[scale=1.0]{fig_6_scattering_scenario.pdf}
  \caption{Scattering from known $\ket{\psi_{\beta}(t,t')}$ state to a constant energy state $\ket{\Psi_{\alpha}(t,t')}$ due to the scattering potential created by impurities in the material.}
  \label{fig:2}
\end{figure}
This phenomenon is illustrated in Fig.~\ref{fig:2}.
We can calculate the scattering amplitude for this scattering scenario using the expression derived in Eq.~(\ref{eq:c14}) as follows
\begin{equation} \label{eq:c16}
  \begin{aligned}
    a_{\alpha\beta}(t,t') =
    -\frac{i}{\hbar}
    \int_{0}^t
    e^{\frac{i}{\hbar}\left(\varepsilon_{\beta} - \varepsilon \right)t_1}
    \bra{\phi_{\beta}(t')}
    V(\vb{r}) \ket{\phi_{\alpha}(t')}  dt_1.
  \end{aligned}
\end{equation}
By assuming for a long duration ($t \rightarrow \infty$), we can turn this integral into a delta distribution
\begin{equation} \label{eq:c17}
  \begin{aligned}
    a_{\alpha\beta}(t') =
    -2\pi i \delta(\varepsilon_{\beta} - \varepsilon)Q,
  \end{aligned}
\end{equation}
where $Q = \bra{\phi_{\beta}(t')} V(\vb{r}) \ket{\phi_{\alpha}(t')}$ and using the completeness properties we can re-write this as
\begin{equation} \label{eq:c18}
    Q =
    \sum_{\vb{k}}\sum_{\vb{k'}}
    \braket{\phi_{\beta}(t')}{\vb{k'}}
    \mel{\vb{k'}}{V(\vb{r})}{\vb{k}}
    \braket{\vb{k}}{\phi_{\alpha}(t')}.
\end{equation}
Moreover, separating $x$-directional and $y$-directional momentum components, we can simply this to obtain
\begin{equation} \label{eq:c19}
    Q =
    \sum_{k_x}\sum_{{k'}_x}
    \int_{-\infty}^{\infty} \int_{-\infty}^{\infty}
    V_{\vb{k'},\vb{k}}
    \phi_{\beta}^{\dagger}(\vb{k'},t')
    \phi_{\alpha}(\vb{k},t')  dk_y d{k'}_y,
\end{equation}
with $V_{\vb{k'},\vb{k}} = \mel{\vb{k'}}{V(\vb{r})}{\vb{k}}$.

@@

Moreover, we can identify the $y$-directional current operator component as
\begin{equation} \label{eq:d12}
  \hat{j}_y = \frac{1}{\widetilde{m}} \left(-i\hbar\pdv{y} - \frac{eE}{\omega}\cos(\omega t) \right).
\end{equation}
Using this operator, we can represent the matrix elements of $y$-directional current operator in Floquet mode basis as
\begin{equation} \label{eq:d13}
  \left({j}_y \right)_{nm,n'm'} =
  \mel{n,k_x}{\;\frac{-1}{\widetilde{m}} \left(i\hbar\pdv{y} + \frac{eE}{\omega}\cos(\omega t) \right)}{n',k'_x}.
\end{equation}
After following the same steps done for the $x$-directional current operator, we can identify the $0$-th component of matrix elements for $y$-directional electric current operator as
\begin{equation} \label{eq:d14}
  \begin{aligned}
    \Big({j}^y_{s=0}\Big)_{nm,n'm'}^{electron} = &
    \frac{ie\hbar}{{m_e}}
    \delta_{k_x,k'_x} \\
    & \times
    \left(
    \sqrt{\frac{n}{2}} \delta_{n',n-1}
    - \sqrt{\frac{n+1}{2}} \delta_{n',n+1}
    \right).
  \end{aligned}
\end{equation}
%===============================================================================
